DATA DRIVEN MODELS FOR APOGEE M DWARFS

### 2018-07-02 paper/draft1/draft1.tex
update figures

\subsection{M dwarf Challenge}
\begin{enumerate}
\item[-] Various attempts / samples for which people have tried to make substantial catalogs of M-dwarfs with physical labels on them (\citealt{West:2011} is the largest M dwarf catalog of $\sim$ 70k sources... Terrien, Mann, Newton, Dressing catalogs)

\item[-] Why is there a problem labeling M dwarfs? Generally, two kinds of approaches to determing physical parameters:

@@

\item[-] Why The Cannon approach? Can be used to amplify a small number of labeled M-dwarfs in a training set into a huge number of labeled M-dwarfs, provided that there are overlapping objects; Make the point that The Cannon does not require us to have good M-dwarf models in the APOGEE wavelength range at all!

\item[-] Describe M dwarf studies that have already been conducted in APOGEE (\citealt{Schmidt:2016} -- analyzed trends and reliablility of ASPCAP parameters for K/M dwarfs; \citealt{Souto:2017} -- modeled 2 M dwarfs, determined Teff/logg/metallicity + 13 elemental abundances; \citealt{Desphande:2013} -- modeled rotational and radial velocites for several hundred M dwarfs; \citealt{Rajpurohit:2018} -- tested BTSettl model grids on 45 M dwarfs, estimated Teff/logg/metallicity)

@@

%using empirical classifications/templates/standards - SPT, luminosity class - this is a relative system but does not provide an absolute standard for measuring physical parameters

%brief importance statment about APOGEE - new instrumentation has allowed for new opportunity - high resolution and large datasets of Mdwarfs are now available

%brief importance statment cannon - better than template matching in that it is using less sparsely sampled known mdwarfs

%ISSUES: models dont accurately reproduce data; empirical standards are relative not absolute

@@

A number of M dwarf studies

@@

The spectral type sample consists of 51 sources, spanning M0-M9 (Figure \ref{fig:train_dist}) cross-matched from the \citealt{West:2011} (hereafter W11) catalog of 78,841 M dwarfs from SDSS. For each source in the catalog, spectral types were determined both through an automated routine using The Hammer \citep{Covey:2007} and by visual inspection to a reported accuracy of $\pm$1 type.

The physical parameter sample consists of 30 sources with reference labels distributed over $3072 < T_{\rm eff} < 4131K$, and -0.48 $<$ [Fe/H] $<$ 0.49 (Figure \ref{fig:train_dist}) drawn from \citealt{Mann:2015} (hereafter M15). The M15 catalog in total contains 183 sources selected from CONCH-SHELL \citep{Gaidos:2013} and with infrared spectra taken in follow up from the SpeX Spectograph. Effective temperatures are measured by comparison of BT-SETTL atmospheric models \citep{Allard:2011} to optical spectra. A subsample of 29 sources with measured angular diameters and bolometric fluxes from long-baseline optical interferometry are used to callibrate effective temperature measurements. Relative uncertainty from the callibration sample is measured to be \color{red}{$\sim$30K (check this)}\color{black}, which added in quadrature to the scatter from model comparison gives a total reported uncertainty of $\sim$60K in T$_{\rm eff}$. Metallicities are measured

@@

\begin{figure}[ht]
\plottwo{figures/west_train_dist.png}{figures/mann_train_dist.png}
\caption{Spectral type model: training label vs. test label.} \label{fig:west_selftest} \label{fig:train_dist}
\end{figure}

@@

APOGEE data products come in several types of files: ap1D, apVisit, apStar, and aspcapStar. SNR of stars in our sample \citealt{Perez:2016}.

@@

\begin{enumerate}
\item[-] A few-paragraph summary of what The Cannon does, with references (\citealt{Ness:2015}, \citealt{Casey:2016}, \citealt{Ho:2017a}, \citealt{Ho:2017b})
\item[-] Describe assumptions of the model: (1) Sources with identical labels have near-identical flux at each pixel; (2) Expected flux at each pixel varies continuosly with change in label. 
\item[-] Briefly describe modeling procedure (training step, test step, quadratic labels, etc.), refer to \citealt{Ness:2015} for full description.
\item[-] Briefly describe model inputs and outputs (1D spectral type model and 2D teff/metallicity and label vectors for those). Mention other models we've tested.
\end{enumerate}

@@

Following from \citealt{Ness:2015} and \citealt{Ho:2017a} 
\begin{equation}
	f_{n\lambda} = g(\ell_n | \theta_{\lambda}) + {\rm noise}
\end{equation}

@@

\begin{figure}[ht]
\plotone{figures/west_spt_self_test.png}
\caption{Spectral type model: training label vs. test label.} \label{fig:west_selftest}
\end{figure}

@@

\subsection{Temperature/Metallicity Model}
\begin{figure}[ht]
\plottwo{figures/mann_teff_self_test.png}{figures/mann_feh_self_test.png}
\caption{Temperature and Metallicity label self test.} \label{fig:mann_selftest}
\end{figure}

\begin{figure}[ht]
\plotone{figures/1_physical_model.png}
\plotone{figures/2_physical_model.png}
\plotone{figures/3_physical_model.png}
\caption{\textit{Top two panels:} Mann-trained model for varying temperatures; \textit{bottom two panels:} Mann-trained model for varying metallicities. \color{red}{Things to add: line list and model - data difference}\color{black}} \label{fig:mann_demo}
\end{figure}

@@

\begin{figure}[ht]
\plotone{figures/west_spt_derivative.png}
\caption{Derivative plots for spectral type model.} \label{fig:west_derivative}
\end{figure}

\begin{figure}[ht]
\plotone{figures/mann_teff_derivative.png}
\plotone{figures/mann_feh_derivative.png}
\caption{Derivative plots for physical parameter model. \textit{Top:} Derivative of flux with repsect to temperature. \textit{Bottom:} Derivative of flux with repsect to metallicity.} \label{fig:mann_derivative}
\end{figure}

@@

>>>>

\begin{enumerate}

\item[-] Define what parameters/classifications are important for Mdwarfs. We want to estimate physical parameters (such as Teff, logg, metallicity) and empirical parameters/classifications (SPT, etc.), which we will refer to as \emph{stellar labels}.

\item[-] Various attempts / samples for which people have tried to make substantial catalogs of M-dwarfs with physical labels on them (\citealt{West:2011} is the largest M dwarf catalog of $\sim$ 70k sources... Terrien, Mann, Newton, Dressing catalogs)

\item[-] brief importance statment about APOGEE - new instrumentation has allowed for new opportunity - high resolution and large datasets of Mdwarfs are now available

\item[-] Describe M dwarf studies that have already been conducted in APOGEE (\citealt{Schmidt:2016} -- analyzed trends and reliablility of ASPCAP parameters for K/M dwarfs; \citealt{Souto:2017} -- modeled 2 M dwarfs, determined Teff/logg/metallicity + 13 elemental abundances; \citealt{Desphande:2013} -- modeled rotational and radial velocites for several hundred M dwarfs; \citealt{Rajpurohit:2018} -- tested BTSettl model grids on 45 M dwarfs, estimated Teff/logg/metallicity). \citealt{Gilhool:2018}; \citealt{Skinner:2018}.

\item[-] Why is there a problem labeling M dwarfs (and especially for high res APOGEE spectra)? Emphasize issues: models dont accurately reproduce data; empirical standards are relative not absolute.

@@

 (3) using empirical classifications/templates/standards - SPT, luminosity class - this is a relative system but does not provide an absolute standard for measuring physical parameters

\item[-] Why The Cannon approach? brief importance statment cannon and what it does (cite \citealt{Ness:2015}, \citealt{Ho:2017a}, \citealt{Casey:2016})

 (1) Can be used to amplify a small number of labeled M-dwarfs in a training set into a huge number of labeled M-dwarfs, provided that there are overlapping objects;

 (2) Make the point that The Cannon does not require us to have good M-dwarf models in the APOGEE wavelength range at all

 (3) better than template matching in that it is using less sparsely sampled known mdwarfs

@@

@@

@@

The spectral type sample consists of 51 sources, spanning M0-M9 cross-matched from the \citealt{West:2011} (hereafter W11) catalog of 78,841 M dwarfs from SDSS. For each source in the catalog, spectral types were determined both through an automated routine using The Hammer \citep{Covey:2007} and by visual inspection to a reported accuracy of $\pm$1 type.

The physical parameter sample consists of 30 sources with reference labels distributed over $3072 < T_{\rm eff} < 4131K$, and -0.48 $<$ [Fe/H] $<$ 0.49 drawn from \citealt{Mann:2015} (hereafter M15). The M15 catalog in total contains 183 sources selected from CONCH-SHELL \citep{Gaidos:2013} and with infrared spectra taken in follow up from the SpeX Spectograph. Effective temperatures are measured by comparison of BT-SETTL atmospheric models \citep{Allard:2011} to optical spectra. A subsample of 29 sources with measured angular diameters and bolometric fluxes from long-baseline optical interferometry are used to callibrate effective temperature measurements. Relative uncertainty from the callibration sample is measured to be \color{red}{$\sim$30K (check this)}\color{black}, which added in quadrature to the scatter from model comparison gives a total reported uncertainty of $\sim$60K in T$_{\rm eff}$.

@@

% \begin{figure}[ht]
% \plottwo{figures/west_train_dist.png}{figures/mann_train_dist.png}
% \caption{Spectral type model: training label vs. test label.} \label{fig:west_selftest} \label{fig:train_dist}
% \end{figure}

@@

We use the aspcapStar spectra (see \citealt{Perez:2016} for a complete description of the pipeline).

APOGEE data products come in several types of files: ap1D, apVisit, apStar, and aspcapStar. SNR of stars in our sample .

@@

\begin{enumerate}
\item[-] Describe the assumptions of the model: (1) Sources with identical labels have near-identical flux at each pixel; (2) Expected flux at each pixel varies continuosly with change in label. 
\item[-] Briefly describe modeling procedure (training step, test step, quadratic labels, etc.), refer to \citealt{Ness:2015} for full description.
\item[-] Briefly describe model inputs and outputs (1D spectral type model and 2D teff/metallicity and label vectors for those). Mention other models we've tested.
\end{enumerate}

@@

Following from the procedure of \citealt{Ness:2015} and \citealt{Ho:2017a} we adopt a model that assumes the flux at each pixel of the spectrum is parameterized as a function \textit{g} 
\begin{equation}
	f_{n\lambda} = g(\ell_n | \theta_{\lambda}) + {\rm noise}
\end{equation}

@@

@@

@@

@@

%%==========================================================================================
%% Figures

\begin{figure}[ht]
\begin{center}
\includegraphics[width=16cm]{figures/west_self_test_spt.png}
\end{center}
\caption{Spectral type model: training label vs. test label.} \label{fig:west_selftest}
\end{figure}

\subsection{Temperature/Metallicity Model}
\begin{figure}[ht]
\begin{center}
\includegraphics[width=16cm]{figures/self_and_validation_test_teff.png}
\includegraphics[width=16cm]{figures/self_and_validation_test_fe_h.png}
\end{center}
\caption{Temperature and Metallicity label self test.} \label{fig:mann_selftest}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=16cm]{figures/demo_derivatives_teff1.png}
\includegraphics[width=16cm]{figures/demo_derivatives_teff2.png}
\includegraphics[width=16cm]{figures/demo_derivatives_teff3.png}
\end{center}
\caption{\textit{Top two panels:} Mann-trained model for varying temperatures.} \label{fig:demo_teff}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=16cm]{figures/demo_derivatives_feh1.png}
\includegraphics[width=16cm]{figures/demo_derivatives_feh2.png}
\includegraphics[width=16cm]{figures/demo_derivatives_feh3.png}
\end{center}
\caption{\textit{Top two panels:} Mann-trained model for varying metallicities.} \label{fig:demo_feh}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=12cm]{figures/west_spt_derivative.png}
\end{center}
\caption{Derivative plots for spectral type model.} \label{fig:west_derivative}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=15cm]{figures/derivative_jackknife_teff.png}
\includegraphics[width=15cm]{figures/derivative_jackknife_feh.png}
\end{center}
\caption{Derivative plots for physical parameter model. \textit{Top:} Derivative of flux with repsect to temperature. \textit{Bottom:} Derivative of flux with repsect to metallicity.} \label{fig:mann_derivative}
\end{figure}
